\Large \bf Dismal Code: Studying the Evolution of Security Bugs

### 2013-09-25 docs/evol.tex
new style - USENIX proceedings

\documentclass{sig-alternate}

@@

\usepackage{ulem}

@@

\title{Dismal Code: Studying the Evolution of Security Bugs}

\numberofauthors{5}

\author{
% 1st. author
\alignauthor
Dimitris Mitropoulos\\
       \affaddr{Department of Management Science and Technology}\\
       \affaddr{Athens University of Economics and Business}\\
       \email{dimitro@aueb.gr}
% 2nd. author
\alignauthor
Vassilios Karakoidas\\
       \affaddr{Department of Management Science and Technology}\\
       \affaddr{Athens University of Economics and Business}\\
       \email{bkarak@aueb.gr}
% 3rd. author
\alignauthor
Panos Louridas\\
       \affaddr{Department of Management Science and Technology}\\
       \affaddr{Athens University of Economics and Business}\\
       \email{louridas@aueb.gr}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Georgios Gousios\\
       \affaddr{Software Engineering Research Group}\\
       \affaddr{Delft University of Technology}\\
       \email{G.Gousios@tudelft.nl}
% 5th. author
\alignauthor Diomidis Spinellis\\
       \affaddr{Department of Management Science and Technology}\\
       \affaddr{Athens University of Economics and Business}\\
       \email{dds@aueb.gr}
}

\maketitle
\begin{abstract}
{\bf Background.} Security bugs are critical programming errors that can lead to serious
vulnerabilities in software. Such bugs may allow an attacker to take over
an application, steal data or prevent the application from working at all.\\
{\bf Aim.} We used the projects stored in the Maven repository to study the
characteristics of security bugs individually and in relation to other software
bugs. Specifically, we studied
the evolution of security bugs through time. In addition, we examined
their persistence and their relationship with
a) the size of the corresponding version, and
b) other bug categories.\\
{\bf Method.} We analyzed every project version of the Maven repository by
using FindBugs, a popular static analysis tool.
To see how security bugs evolve over time we took advantage of
the repository's project history and dependency data.\\
{\bf Results.} Our results indicate that there is no simple rule governing
the number of security bugs as a project evolves.
In particular, we cannot say that across projects security-related
defect counts increase or decrease significantly over time.
Furthermore, security bugs are not eliminated in a
way that is particularly different from the other bugs.
In addition, the relation of security bugs with a project's size
appears to be different from the relation of the bugs
coming from other categories.
Finally, even if bugs seem to have similar behaviour,
severe security bugs seem to be unassociated with other bug categories.\\
{\bf Conclusions.} Our findings indicate that further research should be done
to analyze the evolution of security bugs. Given the fact that our experiment
included only Java projects, similar research could be done for another
ecosystem. Finally, the fact that projects have their own idiosyncrasies
concerning security bugs, could help us find the common characteristics
of the projects where security bugs increase over time.
\end{abstract}

\keywords{Security Bugs, Static Analysis, Software Evolution, Software
Security, Maven, FindBugs.
}
\section{Introduction}

@@

In this paper we present how we used a large software ecosystem to analyse the
relationship of different types of security vulnerabilities to 
evolving software packages.
For our research we used {\it FindBugs},\footnote{\url{http://findbugs.sourceforge.net/}}
a static analysis tool that examines bytecode to detect software bugs and has already been used in
research~\cite{AP10,HP07,SHP06}.
Specifically, we ran FindBugs on all the project
versions of all the projects that exist in the
{\it Maven Central Repository}\footnote{\url{http://mvnrepository.com/}}
(approximately 265GB of data---see Section~\ref{sec:data}).
Then we observed the changes that involved the security bugs and their characteristics.
%This research builds upon our earlier work on the topic~\cite{MGS12}.

@@

Completing the above approaches, our work focuses on the subset of security bugs.
Focusing on such bugs is not a new idea.
Ozment and Schechter~\cite{OS06} examined the code base of the Open{\sc bsd}
operating system to determine whether its security is increasing
over time. In particular, they measured the rate at which new code
has been introduced and the rate at which defects have been reported over
a 7.5 year period and fifteen releases.
Even though the authors present statistical evidence
of a decrease in the rate at which vulnerabilities are being reported,
defects seem to appear persistent for a period of at least 2.6 years.
Massacci et al.~\cite{MNN11} observed
the evolution of software defects by examining six major versions of Firefox.
To achieve this they created a database schema that contained information
coming from the ``Mozilla Firefox-related Security Advisories'' ({\sc mfsa})
list,\footnote{\url{http://www.mozilla.org/projects/security/known-vulnerabilities.html}}
Bugzilla entries and others. Their findings indicated that security bugs are
persistent over time. They also showed that there are many web users that use
old versions of Firefox, meaning that old attacks will continue to work.
Zaman et al.~\cite{ZAH11} focused again on Firefox to study the relation of
security bugs with performance bugs. This was also done by analysing the project's
Bugzilla. Their research presented evidence that security bugs require more experienced developers
to be fixed. In addition, they suggested that security bug fixes are more complex than the
fixes of performance and other bugs.
Shahzad et al.~\cite{SSL12} analysed large sets of vulnerability data-sets to observe
various features of the vulnerabilities that they considered critical. Such features
were the functionality and the criticality of the defects. Their analysis
included the observation of vulnerability disclosures, the behavior of
hackers in releasing exploits for vulnerabilities, patching and others. In
their findings they highlighted the most exploited defects and showed that
the percentage of remotely exploitable vulnerabilities has gradually increased
over the years. Finally, Edwards et al.~\cite{EC12} have recently conducted
a study similar  to ours in which they have considered only four projects.
Their results demonstrate that the number of exploitable bugs does not
always improve with each new release and that the rate of discovery of
exploitable bugs begins to drop three to five years after the
initial release.

@@

Another observation involves bugs that we could call {\it
Security High} and they are a subset of the {\it Security} category.
Such bugs are related to vulnerabilities that appear due to the lack of user-input
validation and can lead to damaging attacks like {\sc sql} injection and
Cross-Site Scripting~\cite{RL12}.
To exploit such vulnerabilities, a malicious user does
not have to know anything about the application internals.
For almost all the other security bugs (coming from
the {\it Malicious Code} category and the rest of
the {\it Security} category bugs),
another program should be written to incorporate references to
mutable objects, access non-final fields, etc.
Also, as bug descriptions indicate,\footnote{\url{http://findbugs.sourceforge.net/bugDescriptions.html}}
if an application has bugs coming from the {\it Security High} category,
it might have more vulnerabilities that FindBugs doesn't report.
Table~\ref{tbl:sev} presents the number
of releases where at least one of these bugs exists. In essence, 5,501 releases
($\approx 4,77\% $), contained at
least one severe security bug. Given the fact that other projects include these
versions as their dependencies, they are automatically rendered vulnerable if
they use the code fragments that include the defects.
The remaining bugs of the {\it Security} category
are grouped together with the bugs of the {\it Malicious Code} category
in another subcategory that we call {\it Security Low}.
This category contains for the most part, bugs
that imply violations of good {\sc oop} (object-oriented programming)
design (i.e. keeping variables private to classes and others).
The above categorization was done specifically to point out the
behaviour of bugs that currently top the corresponding lists of most
security providers.\footnote{\url{http://cwe.mitre.org/top25/index.html}}

@@

We analysed more than 260GB of interdependent project versions to see
how security bugs evolve over time, their persistence, their relation
with other bug categories, and their relationship with size in terms
of bytecode.\footnote{Our data and code are available online at:
\url{https://github.com/bkarak/evol_security_publication_2012}}

@@

\bibliographystyle{abbrv}
\bibliography{evol}

>>>>

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes}
%\documentclass{sig-alternate}

@@

\pdfpagewidth=8.5in
\pdfpageheight=11in

@@

\date{}

\title{\Large \bf Dismal Code: Studying the Evolution of Security Bugs}

\author{
{\rm Dimitris Mitropoulos}\\
Department of Management Science and Technology,
Athens University of Economics and Business
\and
{\rm Vassilios Karakoidas}\\
Department of Management Science and Technology,
Athens University of Economics and Business
\and
{\rm Panos Louridas}\\
Department of Management Science and Technology,
Athens University of Economics and Business
\and
{\rm Georgios Gousios}\\
Software Engineering Research Group,
Delft University of Technology
\and
{\rm Diomidis Spinellis}\\
Department of Management Science and Technology,
Athens University of Economics and Business
}

\maketitle
\thispagestyle{empty}

\subsection*{Abstract}
{\bf Background.} Security bugs are critical programming errors that can lead to serious
vulnerabilities in software. Such bugs may allow an attacker to take over
an application, steal data or prevent the application from working at all.\\
{\bf Aim.} We used the projects stored in the Maven repository to study the
characteristics of security bugs individually and in relation to other software
bugs. Specifically, we studied
the evolution of security bugs through time. In addition, we examined
their persistence and their relationship with
a) the size of the corresponding version, and
b) other bug categories.\\
{\bf Method.} We analyzed every project version of the Maven repository by
using FindBugs, a popular static analysis tool.
To see how security bugs evolve over time we took advantage of
the repository's project history and dependency data.\\
{\bf Results.} Our results indicate that there is no simple rule governing
the number of security bugs as a project evolves.
In particular, we cannot say that across projects security-related
defect counts increase or decrease significantly over time.
Furthermore, security bugs are not eliminated in a
way that is particularly different from the other bugs.
In addition, the relation of security bugs with a project's size
appears to be different from the relation of the bugs
coming from other categories.
Finally, even if bugs seem to have similar behaviour,
severe security bugs seem to be unassociated with other bug categories.\\
{\bf Conclusions.} Our findings indicate that further research should be done
to analyze the evolution of security bugs. Given the fact that our experiment
included only Java projects, similar research could be done for another
ecosystem. Finally, the fact that projects have their own idiosyncrasies
concerning security bugs, could help us find the common characteristics
of the projects where security bugs increase over time.

\section{Introduction}

@@

In this paper we present how we used a large software ecosystem to analyse the
relationship of different types of security vulnerabilities to 
evolving software packages.
For our research we used {\it FindBugs},\endnote{\url{http://findbugs.sourceforge.net/}}
a static analysis tool that examines bytecode to detect software bugs and has already been used in
research~\cite{AP10,HP07,SHP06}.
Specifically, we ran FindBugs on all the project
versions of all the projects that exist in the
{\it Maven Central Repository}\endnote{\url{http://mvnrepository.com/}}
(approximately 265GB of data---see Section~\ref{sec:data}).
Then we observed the changes that involved the security bugs and their characteristics.
%This research builds upon our earlier work on the topic~\cite{MGS12}.

@@

Completing the above approaches, our work focuses on the subset of security bugs.
Focusing on such bugs is not a new idea.
Ozment and Schechter~\cite{OS06} examined the code base of the Open{\sc bsd}
operating system to determine whether its security is increasing
over time. In particular, they measured the rate at which new code
has been introduced and the rate at which defects have been reported over
a 7.5 year period and fifteen releases.
Even though the authors present statistical evidence
of a decrease in the rate at which vulnerabilities are being reported,
defects seem to appear persistent for a period of at least 2.6 years.
Massacci et al.~\cite{MNN11} observed
the evolution of software defects by examining six major versions of Firefox.
To achieve this they created a database schema that contained information
coming from the ``Mozilla Firefox-related Security Advisories'' ({\sc mfsa})
list,\endnote{\url{http://www.mozilla.org/projects/security/known-vulnerabilities.html}}
Bugzilla entries and others. Their findings indicated that security bugs are
persistent over time. They also showed that there are many web users that use
old versions of Firefox, meaning that old attacks will continue to work.
Zaman et al.~\cite{ZAH11} focused again on Firefox to study the relation of
security bugs with performance bugs. This was also done by analysing the project's
Bugzilla. Their research presented evidence that security bugs require more experienced developers
to be fixed. In addition, they suggested that security bug fixes are more complex than the
fixes of performance and other bugs.
Shahzad et al.~\cite{SSL12} analysed large sets of vulnerability data-sets to observe
various features of the vulnerabilities that they considered critical. Such features
were the functionality and the criticality of the defects. Their analysis
included the observation of vulnerability disclosures, the behavior of
hackers in releasing exploits for vulnerabilities, patching and others. In
their findings they highlighted the most exploited defects and showed that
the percentage of remotely exploitable vulnerabilities has gradually increased
over the years. Finally, Edwards et al.~\cite{EC12} have recently conducted
a study similar  to ours in which they have considered only four projects.
Their results demonstrate that the number of exploitable bugs does not
always improve with each new release and that the rate of discovery of
exploitable bugs begins to drop three to five years after the
initial release.

@@

Another observation involves bugs that we could call {\it
Security High} and they are a subset of the {\it Security} category.
Such bugs are related to vulnerabilities that appear due to the lack of user-input
validation and can lead to damaging attacks like {\sc sql} injection and
Cross-Site Scripting~\cite{RL12}.
To exploit such vulnerabilities, a malicious user does
not have to know anything about the application internals.
For almost all the other security bugs (coming from
the {\it Malicious Code} category and the rest of
the {\it Security} category bugs),
another program should be written to incorporate references to
mutable objects, access non-final fields, etc.
Also, as bug descriptions indicate,\endnote{\url{http://findbugs.sourceforge.net/bugDescriptions.html}}
if an application has bugs coming from the {\it Security High} category,
it might have more vulnerabilities that FindBugs doesn't report.
Table~\ref{tbl:sev} presents the number
of releases where at least one of these bugs exists. In essence, 5,501 releases
($\approx 4,77\% $), contained at
least one severe security bug. Given the fact that other projects include these
versions as their dependencies, they are automatically rendered vulnerable if
they use the code fragments that include the defects.
The remaining bugs of the {\it Security} category
are grouped together with the bugs of the {\it Malicious Code} category
in another subcategory that we call {\it Security Low}.
This category contains for the most part, bugs
that imply violations of good {\sc oop} (object-oriented programming)
design (i.e. keeping variables private to classes and others).
The above categorization was done specifically to point out the
behaviour of bugs that currently top the corresponding lists of most
security providers.\endnote{\url{http://cwe.mitre.org/top25/index.html}}

@@

We analysed more than 260GB of interdependent project versions to see
how security bugs evolve over time, their persistence, their relation
with other bug categories, and their relationship with size in terms
of bytecode.\endnote{Our data and code are available online at:
\url{https://github.com/bkarak/evol_security_publication_2012}}

@@

{\footnotesize \bibliographystyle{acm}
\bibliography{evol}
